Seasonality of submesoscale dynamics in the Kuroshio Extension

### 2016-09-18 writeup/RochatEtal_ResponseToReviewers.tex

\maketitle

\section*{Summary}

In this revised manuscript, we have addressed the reviewers'
comments. The most fundamental change is the inclusion of a wavenumber-frequency
spectral analysis suggested by reviewer 3. The new results vindicate
our claim that inertia-gravity waves, particularly super-inertial waves,
account for most of the summertime SSH variance and surface KE at scales
between 10 and 100 km. We have also expanded the introductions and discussion
to place our study in the context of Surface Water \& Ocean Topography (SWOT)
satellite mission --- we discuss
the implications of our findings for the estimation of surface velocity from
high-resolution sea-surface height.

We thank the reviewers  for the speedy review, constructive
criticism, and thoughtful suggestions that have helped us improve our discussion and
forced us to provide a better context for our study. We hope our manuscript is
now suitable for publication and look forward to your reply.

@@

We thank reviewer \#1 for the speedy review. Below we address the reviewer’s
comments and suggestions.

\begin{enumerate}

\item {\it In the bottom paragraph on P.10, the authors emphasized that "At scales larger
than 20km, the April and October spectra based on hourly velocity snapshots are
nearly indistinguishable from each other". This emphasis appears to be contradictory
to the statements in other places of the manuscript that the seasonality of subinertial
submesoscale flows is strong (e.g., those relating to Figure 1b-c and Figure 2).}\\

    \bdp
        We agree that the emphasis on the similarity of hourly spectra can appear
        contradictory. Those spectra (now Figure 4c
        and supporting information) are similar because both submesoscale sub-inertial
        and super-inertial flows undergo a strong seasonal cycle that is out of
        phase --- there is a phase cancellation and the spectra based on hourly
        data are indistinguishable within errorbars. In any event, we removed
        that sentence from the text. The new Figure 4 should convey
        the phase cancellation idea more effectively.
    \edp

@@

\subsection{Major points}

\begin{enumerate}

  \item {\it First of all, generation mechanism of submesoscale inertial-gravity waves in summer
        and early fall is unclear to me. In the paper, authors attribute the mechanism to
        internal tides (X-12, section 6, last sentence) and emphasize using the tidal forcing in
        their model (X-5, section 2, 2nd paragraph). However, authors do not explain or show
        any interaction mechanism of internal tides (please at least add some reference). With
        the fixed frequencies of tidal constituents and existence of the critical latitude for the
        diurnal tides, I am not sure what mechanism can represent the surface near-inertial
        wave generation within the whole analyzed region by internal tides (with the
        horizontal scales of 10-100km). If authors compare Figures 2 and 4 with those
        estimated from the non-tidal forcing run in the Supplemental material, authors'
        conclusion could be strengthened.}\\

        \bdp
          The generation mechanism for internal
          waves involves changes in the wind (NIWs) and barotropic tides sloshing over
          topographic features --- we are not claiming that there exists a different
          generation mechanism
          in summer compared to winter; neither are we claiming that surface NIWs
          are generated by internal tides. What is different is the near-surface
          expression of IGWs, which is very sensitive to the near-surface stratification
          (D'saro 1978; cf. second paragraph of discussion).\\

@@

        \bdp
            We are now citing the recent review on NIWs by \cite{alford_etal2016} that contains
            a section dedicated to mesoscale-turbulence-NIW interactions ---  \cite{alford_etal2016}
            cite all the
            aforementioned papers. We believe the new discussion together with Figure 4
            and Figure S7 will help clear any ambiguity. Those figures unambiguously
            show that most of the IGWs with horizontal scales between 10-100 km
            are mostly super-inertial (not near-inertial); the refracted/dispersed  NIWs
            propagate into the interior and only weakly project onto SSH \citep{alford_etal2016}.
        \edp

  \item {\it Finally, I don't understand why authors define 10-100 km as a submesoscale and used
        this scale for the spatial filtering (X-7, section 3). Authors should discuss Figure 4 first
        and then give a reason why they define 10-100 km as a submesoscale and use this
        scale for the filtering. In X-11, section 5, author point out that a part of the dynamics
        around 250 km scale should be explained by near-inertial waves. Then, I wonder why
        authors decide to discuss internal waves in 10-100 km scale.}

        \bdp
            Following reviewer \#3's suggestion, we have added paragraphs about SWOT
            to the introduction and conclusion. This discussion adds context to our study
            and justifies the focus on the horizontal scales between 10-100 km. Our definition
            of submesoscales have pros and cons (see reply to reviewer \#3's
            minor comment \#1) but is consistent with the definition used by
            (at least part of) the SWOT community (Fu\& Ferrari 2008). Giving the
            definition, which we present in the first sentence of the introduction,
            then there is no ambiguity in using 100 km as a cut-off scale for the filter.
            We have chosen to present bulk statistics first, and then discuss details of
            different scales --- this makes more sense now with the new wavenumber-frequency
            analysis.
        \edp

@@

        \bdp
            We are aware of the this issue --- the effective resolution based on the KE spectrum
            was already mentioned
            in the supporting information to the first draft.  In
            fact, the smallest scale discussed in our paper (10 km) was set by the effective
            resolution of highest-resolution simulation. To be more explicit, we have now included
            the effective resolution together with the nominal resolutions in the main text (Section 2).\\

            The ``unresolved energy'' does not come from anywhere --- flows smaller than
            about 10 km are strongly damped.
            To the extent that 1-km-scale flows are associated with a forward
            energy transfer, then their effects on the scales concerning our study
            (10-100 km) are likely energetically insignificant.
            Also important to this discussion is that higher resolution simulations will resolve {\it both}
            1-km-scale eddies and 1-km-scale IGWs that are generated via wave-wave
            interactions and wave-vorticity interactions.\\

            \cite{buckingham_etal2016} made a very good point that models and observations
            to date may not resolve the mixed-layer deformation scale in summer; this is no
            different for our simulations. But the interpretation that this is the main
            source of seasonally varying eddy kinetic energy is unclear because
            summer is also the season when lateral buoyancy
            gradients (and thus available potential energy) are minimum within the mixed layer
            \citep[e.g., ][]{callies_etal2015}.\\

@@

        One way to define mesoscale and submesoscale is to define what scales of motion a
        particular sensor will resolve. While this would not be optimal, it is objective. This is
        probably the motivation for the definition used by Callies and Ferrari (2013), since the
        swath altimeter (as initially proposed) would resolve scales greater than 1 km but
        smaller than the present-day, profiling satellite altimeters – i.e., eddy-like features
        with diameters greater than 80-100 km (see Chelton et al. 2011, Progress).
        Unfortunately, this definition has the adverse effect of changing for each sensor.

        A better definition of the submesoscale might be founded on dynamical arguments.
        Wunsch and Stammer (1998) introduce the geostrophic equations and note that the
        lower limit for which this approximation is strictly valid is ~30 km. This conveniently
        defines the lower limit of the mesoscale regime. I would therefore define the
        submesoscale regime as anything smaller than this but one for which the hydrostatic
        approximation remains valid (e.g., 1 km).

      The motivation for wanting to use the term ``submesoscale'' in the title is clear –
      submesoscale motions are a hot topic within the community and it catches individuals’
      eyes. This is fine. But what must be made clear within the manuscript is what scales of
      motion and what force balances (i.e., dynamical regimes) are being described.}\\

      \bdp
          We respect the reviewer's opinion about the definition of submesoscales.
          But there seems to be no clear definition in the literature --- the
          Stammer $\&$ Wunsch paper states that their
          definition of mesoscales is ``very rough'' and ``there is no generally
          agreed upon definition''. There is also irritation that oceanographers
          decided to term mesoscales the flows with dynamics analogous to
          the meteorological synoptic scale motions \citep[Appendix C of ][]{wunsch2015}.\\

@@

          Some investigators define mesoscales as flows with scales near the
          1st baroclinic deformation radius. But there is significant confusion
          owing to factors of  $2 \pi$ --- e.g., some investigators would use $R_1 = 30$ km,
          others would argue for $L_1 = 2\pi \times 30 \approx 190$ km
          \citep{wunsch_ferrari2004,ferrari_wunsch2009}.
          To add yet another controversy, \cite{larichev_held1995} argue
          that the deformation length is not the panacea that linear theorists
          suggest. In baroclinic geostrophic turbulence, the bulk of the
          energy production occurs on scales much larger than the deformation
          radius; the energy-containing scales of equilibrated baroclinic geostrophic
          turbulence are also larger than the deformation radius \cite{larichev_held1995}.\\

          According
          to our definition the mesoscales are the energy-containing scales of the
          flow --- the eddies that contain most of the eddy kinetic energy in the ocean
          \citep{ferrari_wunsch2009}. From the the centroid of KE wavenumber
          spectrum from current altimeters, this scale varies roughly from 100-300 km.
          We then \textit{define} submesoscales as
          the energetically subdominant scales, the scales smaller than about 100 km.
          Our lower wavelength definition is limited by the effective resolution
          of our higher resolution simulation ($\sim$10 km).
          This definition is consistent with \cite{callies_ferrari2013} (they did not
          define mesoscales/submesoscales based on scales that a sensor resolves
          as the reviewer suggests).\\

          We understand that our definition is imperfect --- we may satisfy some readers but
          irritate others. But
          besides the unsettled definitions in our community and consistency with our recent
          work \citep{rocha_etal2016}, there are at least two reasons for sticking to
          our definition: (i) this is the definition used by the SWOT community (Fu \& Ferrari
              2008); incidentally current altimeters do not resolve
              scales smaller then about 100 km; and (ii) our goal is to determine which flows
              (geostrophically balanced turbulence, inertia-gravity waves, etc)
              dominate the 10-100 km horizontal
              scales and their seasonality. In any event,  readers of the first sentence of our
              abstract will find our explicit definition of submesoscales.
      \edp

@@

  \item {\it (Section 1, Paragraph 2) Model resolutions: it would be helpful for readers to place in
        parentheses nominal horizontal resolutions corresponding to these spatial resolutions.}

        \bdp
          We have added both nominal and effective resolutions.
        \edp

  \item {\it (Section 1, Paragraph 3) What types of horizontal and temporal scales for IGWs are
        we talking about here? How would this fall onto the Garrett-Munk (GM) internal
        wave spectrum? (Recall: the GM spectrum contains both balanced and unbalanced
        flows, not just internal waves, despite that the name suggests this.)}\\

        \bdp
          We explicitly state that those are submesoscale IGWs. Using
          our definition of submesoscales, these are IGWs
          with horizontal scales about 10-100 km\footnote{Bill Young suggested the
          term submesoscale IGWs.}; IGWs have periods between the inertial and buoyancy
          periods.  The GM spectrum is inaccurate
          close to the surface (most of our results concern surface currents and SSH,
          and it does not account for tidal and inertial peaks. We believe that the
          new plot of the wavenumber-frequency spectrum of the SSH
          variance (new Figure 4) --- following the reviewer's suggestion ---
          will more clearly convey the idea that there is no horizontal scale
          separation between submesoscale IGWs and submesoscale turbulence.\\

          If the reviewer is using ``balanced'' as a shorthand to ``geostrophically
          and hydrostatically
          balanced'', then the comment about the GM spectrum containing both balanced
          and unbalanced flows is incorrect.  GM used linear theory
          of IGW to synthesize observations, mostly mooring data, in the IGW band
           --- they
          fitted an analytic model to (super-inertial) frequency spectra and
          different vertical modes (see Walter Munk's chapter
          in \textit{Evolution of Physical Oceanography} 1981 and references therein).
          One could argue
          that other flows project onto similar temporal scales, but those motions
          are super-inertial (e.g., stratified turbulence), not (sub-inertial)
          balanced flows.
        \edp

@@

\item {\it (Section 5, Paragraph 3) Remove the exclamation mark.}\\

    Removed.

\item {\it (Section 4, first paragraph) I understand the motivation for examining the Laplacian of
      sea surface height but the average reader will not understand the connection.}

      \bdp
      Readers familiar with basic geostrophy concepts are likely to understand the connection.
      We make the connection between the Laplacian of SSH and relative vorticity
      in the caption of figure 3.
      \edp

@@

\item {\it (Figure 1) ** Place panels (d) and (e) before panels (b) and (c) since this is the order
        in which you refer to them in the manuscript. ** Difficult to see contours in (d) and
        (e). ** Consider overlaying mixed layer depth on these transects since it is mentioned
        in the text.}

        \bdp
          We have changed the order of the panels and plotted the mixed layer depth
          on the density plots.
        \edp

\item {\it (Figure 2) Is the mean of the strain rate zero? Is the mean of the divergence zero? The
        variance should be used rather than the root-mean-squared (RMS).}

        \bdp
            At every snapshot, the horizontally-averaged divergence is nearly zero;
            the strain rate horizontal average is not zero, because by definition
            $\alpha \ge 0$. The (normalized) RMS better characterizes the the bulk
            gradient Rossby number of the flow.
        \edp

@@

      {\it ``Any horizontal velocity field, uh, can be expressed in terms of vorticity, divergence,
      strain rate and mean flow. This can be seen by expressing uh as a Taylor series
      expansion and then decomposing the velocity gradient into symmetric and anti-
      symmetric parts [Landau and Lifshitz, 1987]. Because vortices, vertical motion and
      fronts are ubiquitous at the submesoscale, this is a natural decomposition and
      variances of these quantities are apt descriptors of submesoscale turbulence.''}\\

      \bdp
        Adding the elementary motivation above is yet another controversy: some
        readers may appreciate it but others may find it condescending.
        We argue that most physical oceanographers  will likely
        likely know this material. Even readers from other areas, who did not take
        basic fluid mechanics, will have the intuition that if the flow peaks
        at mesoscales (energy-containing), then second-order statistics of the
        velocity gradient will better highlight the submesoscales than the
        KE (second-order statistic of the velocity).
      \edp

\item {\it (Figure 3) Would be nice to include a third row that contains the collapsed (i.e., one-
      dimensional) PDFs of normalized vorticity to illustrate the skewness of hourly and
      daily-averaged flow. I believe you already have this information in the supplementary
      material but would be nice to illustrate to the reader. Also, in your normalization of
      vorticity, are you accounting for the fact that the Coriolis parameter changes with
      latitude? Your domain spans considerable changes in f and the relevant dynamical
      parameter is the gradient Rossby number.}

      \bdp
        We have experimented with a new figure 3, including a third row to show the
        vorticity PDFs. But the extra plots add little extra, if any, information
        and take a lot of space. For what it is worth, we explicitly give the relevant statistic
        (skewness) for each month, etc.  As the reviewer mentions, the collapsed
        PDFs for vorticity and divergence are shown in Figure S2, which compares
        April/October and hourly/daily-averaged in a single plot.\\

@@

        Yes, as we mention in the text, we have normalized by those
        quantities by the \textit{local} inertial frequency, not the inertial
        frequency at mid-latitude.
      \edp

\item {\it (Figure 4b) Strictly speaking, the label on the y-axis should not be SSH variance. SSH
      variance would be the integral of this curve. The y-axis should be labelled as the
      power spectral density of sea surface height (SSH), or SSH Spectra.}\\

      \bdp
      We label the y-axis of the SSH variance spectrum as \textit{SSH variance density}
      --- short for \textit{SSH variance spectral density} --- not \textit{SSH variance} as the reviewer
      suggests.  We chose to use the short term to avoid cluttering the label
      since there is no ambiguity because we include the
      units of SSH variance per unit wavenumber.
      We further disagree that the y-axis should be labelled \textit{power spectral density of SSH}
      or \textit{SSH spectra}. On physical grounds, \textit{power} adds
      unnecessary ambiguity,
      because SSH variance density has nothing to do with \textit{power}.\footnote{The
      term \textit{power spectral density}  appears to have been introduced
      for \textit{frequency} spectrum of various quantities in signal processing (Rob Pinkel, personal
      communication).}
      And ``SSH [wavenumber] spectrum'' is the plot of SSH variance spectral density
      as a function of wavenumber, not the quantity in the y-axis.
      \edp

### 2016-09-18 writeup/rocha_etal.tex
Minor edits by Sarah and Teri.

% now SWOT
The characterization of the partition between geostrophic and ageostrophic flows across
the submesoscales and their seasonality has immediate applications for the planning
of the Surface and Water Ocean Topography (SWOT) satellite mission.
SWOT will provide
sea-surface height (SSH) with about 15 km resolution, thereby providing the
first global SSH measurements at submesoscales \citep{fu_ubelmann2014}. A
challenge for SWOT is the presence of ageostrophic flows that straddle the submesoscales and project
onto SSH \citep[e.g., ][]{richman_etal2012} --- the
nuisance\footnote{``One man's noise is
another man's signal'' (Walter Munk).} may upset the estimation
of submesoscale surface geostrophic velocities.

@@

\section{The LLC numerical simulations}
We use results of two latitude-longitude polar cap (LLC)
realistic numerical simulations. The outputs
analyzed here, LLC2160 (nominal resolution 1/24$^\circ$$\approx$4.7 km;
effective resolution $\approx$ 20 km) and LLC4320
(nominal resolution 1/48$^\circ$$\approx$2.3 km;
effective resolution $\approx$8 km),   are
forward Massachusetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
numerical solutions on a LLC grid  \citep{forget_etal2015} with
90 vertical levels. The coarser-resolution LLC simulation was
spun up from an Estimating the Circulation and Climate of the Ocean, Phase II \citep[ECCO2; ][]{menemenlis_etal2008}
adjoint-method state estimate, constrained to millions
of observations from 2009 through 2011. Both simulations were forced by
tides hourly and by 6-hourly surface atmospheric fields. The LLC2160
output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
the LLC2160 simulation, spanning one year from September 2011 to October 2012.
The LLC4320 simulation is an extension of the 3-month long output used by
\citet{rocha_etal2016} --- for details, see supporting information.

@@

The seasonal cycle of the horizontal divergence, however, showcases the complexity
of the upper-ocean annual variability. If submesoscale eddies and fronts dominated
the near-surface variability all year, then the seasonal cycle of horizontal divergence,
vertical vorticity, and lateral strain rate would be in phase \citep[e.g.,][]{sasaki_etal2014}.
While there is a clear wintertime peak in divergence of daily-averaged velocity
(see dashed lines in figure \ref{fig2}c; RMS divergence $\sim0.1 f$ in the 1/48$^\circ$
simulation), the hourly fields show
a stronger enhancement of
lateral divergence in late summer/early fall (RMS divergence $\sim0.22 f$ the 1/48$^\circ$
simulation). Because the 1/48$^\circ$ simulation better resolves
smaller-scale submesoscale flows, a secondary RMS divergence peak in winter  is
nearly as strong as in summer. Submesoscale fronts and eddies evolve
relatively fast, and there is no clear
temporal and spatial scale separation between those motions and IGWs
\citep[][]{mcwilliams2016}:
daily-averaging the velocity fields efficiently suppresses the summertime horizontally
divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
Figure \ref{fig2}c also shows that most of the lateral divergence is associated
with submesoscale flows: smoothing the velocity fields with a 100-km-cutoff
suppresses more than 80$\%$ of the RMS divergence.

@@

The April vorticity-strain joint-PDF has a shape characteristic of submesoscale turbulence
 (Figures \ref{fig3}a-b). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
 with strong positive skewness are fingerprints of submesoscale fronts
 \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
 joint-PDF is similar for hourly and daily-averaged fields, although
 the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
 The April results are characteristic of winter, indicating that wintertime submesoscale
 surface velocity is strongly dominated by submesoscale turbulence.
 The hourly velocity and sea-surface heigh fields have an important
ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
winter (Figure \ref{fig3}e). The daily-averaged fields
are largely
in  geostrophic balance (Figure \ref{fig3}f).

The October vorticity-strain joint-PDF shows much weaker skewness (the vorticity skewness
is 0.68 and 0.67 for hourly and daily-averaged velocities). The shape of the
vorticity-strain joint-PDF  appears to be a combination of two half-ellipses centered
about $\zeta=0$, one with a 45$^\circ$ slope (characteristic of submesocale fronts
that persist in summer) and one with a very steep slope.
That the submesoscale dynamics in October are mainly ageostrophic is clearly depicted
in the shape of joint-PDF of vorticity-Laplacian of sea-surface height for hourly fields
 (Figure \ref{fig3}g), which is an ellipse aligned in the vertical axis.
Daily-averaging the model suppresses the ageostrophic, super-inertial flows, and, therefore,
the daily-averaged flow is essentially geostrophic as depicted by the 45$^\circ$-tilted
ellipse in the vorticity-Laplacian of sea-surface height joint-PDF.

 Time series of PDFs of vorticity and divergence (supporting information) show
 a strong oscillation between these two regimes. In late winter/early spring
 the vorticity is strongly positively skewed, whereas the divergence is moderately
 negatively skewed (convergence/downwelling) as predicted by frontogenesis \citep[e.g., ][]{mcwilliams2016}
. In late summer/early fall, the divergence is stronger, but PDFs are much less skewed,
consistent with linear IGWs.

\section{Wavenumber-frequency spectrum}
To confirm that submesoscale IGWs account for most of the super-inertial SSH variance
and surface KE, we calculate wavenumber-frequency spectrum of SSH variance. Focusing on
the high-frequency
content, we compute the wavenumber-frequency spectrum  every 10 days and average
the results to obtain a spectral estimate. Before calculating the discrete Fourier
transform, we
remove linear trends and multiply the data by a three-dimensional Hanning window.
We azimuthally-average the spectrum in wavenumber.

Figures \ref{fig4}a-b show the wavenumber-frequency of SSH variance for the LLC4320
simulation. Mesoscales flows and the lunar semi-diurnal tide contain most of the
SSH variance. But there is significant SSH variance at submesoscales, both sub-inertial
and super-inertial. Sub-inertial SSH variance at submesoscales is larger in April, whereas
super-inertial SSH variance at submesoscales is larger in October. At scales smaller
than 100 km (submesoscales), there the SSH variance is 66$\%$ larger in October than in April.
IGWs, particularly modes two through four, account for most of the SSH variance
increase in summer --- higher vertical
modes are more sensitive to changes in near-surface stratification (supporting information).
In October, super-inertial IGWs account for a SSH standard deviation of about 4.9 cm,
comparable to the April sub-inertial SSH standard deviation of $\sim$5.2 cm (Table \ref{tab1}).
The
Similarly, the submesoscale super-inertial surface KE is 63$\%$ larger in October
than the April (supporting information).

@@

Figure \ref{fig4}c shows the integral over frequency
of wavenumber-frequency spectra in figures \ref{fig4}a-b. ,
Both in April and October, the
SSH wavenumber spectrum approximately follows a -2 power-law at submesoscales
(solid lines in figure \ref{fig4}c)  --- there is
small difference between the spectra owing to the phase cancellation between the
seasonal cycle of submesoscale turbulence and inertia-gravity waves. The sub-inertial flow
has steeper submesoscale SSH variance wavenumber spectrum (dashed lines in figure \ref{fig4}c).
 The difference between the SSH variance wavenumber spectrum for all frequencies and
 sub-inertial frequencies is dramatic in October when super-inertial flows account for
 about 80\% of the SSH variance and KE at submesoscales. We obtain similar results by computing
SSH variance and KE wavenumber spectra directly from hourly and daily-averaged
SSH velocity fields.

@@

Internal tide energy flux does not show a seasonal modulation \citep[e.g.,][]{alford2003},
hence one expects a strong seasonality in the near-surface
expression of internal tides and
higher frequency IGWs if the near-surface stratification varies strongly.
Windy near-inertial wave (NIW) generation is bursty and peaks in early winter, but
NIW energy is similar energy level in October and April
\citep{alford_etal2016}. Furthermore, in our simulations the higher frequency inertia-gravity-wave
account for most of the seasonality in super-inertial surface KE and SSH
(Figure \ref{fig4}a-b),

@@

\begin{acknowledgments}
 William R. Young and the reviewers --- Christian Buckingham and two anonymous referees ---
  provided exceptionally helpful and constructive comments. Greg Wagner
 suggested that the near-surface shape of the baroclinic modes changes seasonally.
 We thank the MITgcm community and our colleagues at the NASA Advanced
Supercomputing (NAS) Division for their awesome support.
This research was funded by NSF (OCE1357047) and NASA (NNX13AE44G,NNX13AE85G,NNX16AH67G).
The LLC output can be obtained from the ECCO project (\texttt{http://ecco2.org/llc$\_$hires}). The altimeter products were produced by Ssalto/Duacs
and distributed by AVISO, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).
Codes and output files are available online at the project repository
 (\texttt{https://github.com/crocha700/UpperOceanSeasonality}).
\end{acknowledgments}

>>>>

% now SWOT
The characterization of the partition between geostrophic and ageostrophic flows across
the submesoscales and their seasonality has immediate applications for the planning
of the Surface and Water Ocean Topography (SWOT) satellite mission.
SWOT will provide
sea-surface height (SSH) with about 15 km resolution, thereby providing the
first global SSH measurements at submesoscales \citep{fu_ubelmann2014}. A
challenge for SWOT is the presence of ageostrophic flows that straddle the submesoscales and project
onto SSH \citep[e.g., ][]{richman_etal2012} thus making submesoscale geostrophic
velocities difficult to extract from SSH.

@@

\section{The LLC numerical simulations}
We use results of two latitude-longitude polar cap (LLC)
realistic numerical simulations. The outputs
analyzed here, LLC2160 (nominal resolution 1/24$^\circ$$\approx$4.7 km;
effective resolution $\approx$ 20 km) and LLC4320
(nominal resolution 1/48$^\circ$$\approx$2.3 km;
effective resolution $\approx$8 km),   are
forward solution of the Massachussetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
on a LLC grid  \citep{forget_etal2015} with
90 vertical levels. The coarser-resolution LLC simulation was
spun up from an Estimating the Circulation and Climate of the Ocean, Phase II \citep[ECCO2; ][]{menemenlis_etal2008}
adjoint-method state estimate, constrained to millions
of observations from 2009 through 2011. Both simulations were forced by
tides hourly and by 6-hourly surface atmospheric fields. The LLC2160
output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
the LLC2160 simulation, spanning one year from September 2011 to October 2012.
The LLC4320 simulation is an extension of the 3-month long output used by
\citet{rocha_etal2016} --- for details, see supporting information.

@@

The seasonal cycle of the horizontal divergence, however, showcases the complexity
of the upper-ocean annual variability. If submesoscale eddies and fronts dominated
the near-surface variability all year, then the seasonal cycle of horizontal divergence,
vertical vorticity, and lateral strain rate would be in phase \citep[e.g.,][]{sasaki_etal2014}.
While there is a clear wintertime peak in divergence of daily-averaged velocity
(see dashed lines in figure \ref{fig2}c; RMS divergence $\sim0.1 f$ in the 1/48$^\circ$
simulation), the hourly fields show
a stronger enhancement of
lateral divergence in late summer/early fall (RMS divergence $\sim0.22 f$ in the 1/48$^\circ$
simulation). Because the 1/48$^\circ$ simulation better resolves
smaller-scale submesoscale flows, a secondary RMS divergence peak in winter  is
nearly as strong as in summer. Submesoscale fronts and eddies evolve
relatively fast, and there is no clear
temporal and spatial scale separation between those motions and IGWs
\citep[][]{mcwilliams2016}:
daily-averaging the velocity fields efficiently suppresses the summertime horizontally
divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
Figure \ref{fig2}c also shows that most of the lateral divergence is associated
with submesoscale flows: smoothing the velocity fields with a 100-km-cutoff
suppresses more than 80$\%$ of the RMS divergence.

@@

The April vorticity-strain joint-PDF has a shape characteristic of submesoscale turbulence
 (Figures \ref{fig3}a-b). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
 with strong positive skewness are fingerprints of submesoscale fronts
 \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
 joint-PDF is similar for hourly and daily-averaged fields, although
 the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
 The April results indicate that wintertime submesoscale
 surface velocity is strongly dominated by submesoscale turbulence.
 The hourly velocity and sea-surface height fields have an important
ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
winter (Figure \ref{fig3}e). The daily-averaged fields
are largely
in  geostrophic balance (Figure \ref{fig3}f).

The October vorticity-strain joint-PDF shows much weaker skewness (the vorticity skewness
is 0.68 and 0.67 for hourly and daily-averaged velocities). The shape of the
vorticity-strain joint-PDF  appears to be a combination of two half-ellipses centered
about $\zeta=0$, one with a 45$^\circ$ slope (characteristic of submesocale fronts
that persist in summer) and one with a very steep slope.
That the submesoscale dynamics in October are mainly ageostrophic is clearly depicted
in the shape of joint-PDF of vorticity-Laplacian of sea-surface height for hourly fields
 (Figure \ref{fig3}g), which is an ellipse aligned in the vertical axis.
Daily-averaging the model suppresses the ageostrophic, super-inertial flows, and, therefore,
the daily-averaged flow is essentially geostrophic as depicted by the 45$^\circ$-tilted
ellipse in the vorticity-Laplacian of sea-surface height joint-PDF (Figure \ref{fig3}h).

 Time series of PDFs of vorticity and divergence (supporting information) show
 a strong oscillation between these two regimes. In late winter/early spring
 the vorticity is strongly positively skewed, whereas the divergence is moderately
 negatively skewed (convergence/downwelling) as predicted by frontogenesis \citep[e.g., ][]{mcwilliams2016}.
 In late summer/early fall, the divergence is stronger, but PDFs are much less skewed,
consistent with linear IGWs.

\section{Wavenumber-frequency spectrum}
To confirm that submesoscale IGWs account for most of the super-inertial SSH variance
and surface KE, we calculate the wavenumber-frequency spectrum of SSH variance. Focusing on
the high-frequency
content, we compute the wavenumber-frequency spectrum  every 10 days and average
the results to obtain a spectral estimate. Before calculating the discrete Fourier
transform, we
remove linear trends and multiply the data by a three-dimensional Hanning window.
We azimuthally-average the spectrum in wavenumber.

Figures \ref{fig4}a-b show the wavenumber-frequency of SSH variance for the LLC4320
simulation. Mesoscales flows and the lunar semi-diurnal tide contain most of the
SSH variance. But there is significant SSH variance at submesoscales, both sub-inertial
and super-inertial. Sub-inertial SSH variance at submesoscales is larger in April, whereas
super-inertial SSH variance at submesoscales is larger in October. At scales smaller
than 100 km (submesoscales), the SSH variance is 66$\%$ larger in October than in April.
IGWs, particularly modes two through four, account for most of the SSH variance
increase in summer --- higher vertical
modes are more sensitive to changes in near-surface stratification (supporting information).
In October, super-inertial IGWs account for a SSH standard deviation of about 4.9 cm,
comparable to the April sub-inertial SSH standard deviation of $\sim$5.2 cm (Table \ref{tab1}).
The
Similarly, the submesoscale super-inertial surface KE is 63$\%$ larger in October
than in April (supporting information).

@@

Figure \ref{fig4}c shows the integral over frequency
of wavenumber-frequency spectra in figures \ref{fig4}a-b.
Both in April and October, the
SSH wavenumber spectrum approximately follows a -2 power-law at submesoscales
(solid lines in figure \ref{fig4}c)  --- there is
small difference between the spectra owing to the phase cancellation between the
seasonal cycle of submesoscale turbulence and inertia-gravity waves. The sub-inertial flow
has steeper submesoscale SSH variance wavenumber spectrum (dashed lines in figure \ref{fig4}c).
 The difference between the SSH variance wavenumber spectrum for all frequencies and
 sub-inertial frequencies is dramatic in October when super-inertial flows account for
 about 80\% of the SSH variance and KE at submesoscales. We obtain similar results by computing
SSH variance and KE wavenumber spectra directly from hourly and daily-averaged
SSH velocity fields.

@@

Internal tide energy flux does not show a seasonal modulation \citep[e.g.,][]{alford2003};
hence one expects a strong seasonality in the near-surface
expression of internal tides and
higher frequency IGWs if the near-surface stratification varies strongly.
Windy near-inertial wave (NIW) generation is intermittent and peaks in early winter, but
NIW energy has a similar energy level in October and April
\citep{alford_etal2016}. Furthermore, in our simulations the higher frequency inertia-gravity-wave
account for most of the seasonality in super-inertial surface KE and SSH
(Figure \ref{fig4}a-b),

@@

\begin{acknowledgments}
 William R. Young and the reviewers --- Christian Buckingham and two anonymous referees ---
  provided exceptionally helpful and constructive comments. Greg Wagner
 suggested that the near-surface shape of the baroclinic modes changes seasonally.
 We thank the MITgcm community and our colleagues at the NASA Advanced
Supercomputing (NAS) Division for their awesome support.
This research was funded by NSF (OCE1357047) and NASA (NNX13AE44G,NNX13AE85G,NNX16AH67G,NNX16AO5OH).
The LLC output can be obtained from the ECCO project (\texttt{http://ecco2.org/llc$\_$hires}). The altimeter products were produced by Ssalto/Duacs
and distributed by AVISO, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).
Codes and output files are available online at the project repository
 (\texttt{https://github.com/crocha700/UpperOceanSeasonality}).
\end{acknowledgments}

### 2016-09-18 writeup/rocha_etal_supporting-info.tex
Minor edits by Sarah and Teri.

Table \ref{tab:llc} shows details of the LLC spin-up hierarchy.
Both simulations analyzed in the present paper, LLC2160 and LLC4320,
were forced by surface fluxes from the 0.14$^\circ$ European Centre for
Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
 starting in 2011, and the 16 most significant  tidal components.
The LLC code and model configuration files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}). The model output can be obtained from  obtained from the ECCO project (\texttt{http://ecco2.org/llc$\_$hires}).

@@

To assess how well the LLC simulations represents the high-frequency variability,
we compare the model near-surface velocity field against two available
mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
For proper comparison, we use model data closest to these two moorings (west of
the region considered in the present study). Focusing on high-frequencies, we
calculate frequency spectral estimates every month. There are about 48 independent
spectral realizations.

>>>>

Table 1 shows details of the LLC spin-up hierarchy.
Both simulations analyzed in the present paper, LLC2160 and LLC4320,
were forced by surface fluxes from the 0.14$^\circ$ European Centre for
Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
 starting in 2011, and the 16 most significant  tidal components.
The LLC code and model configuration files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}). The model output can be obtained from  obtained from the ECCO project (\texttt{http://ecco2.org/llc$\_$hires}).

@@

To assess how well the LLC simulations represents the high-frequency variability,
we compare the model near-surface velocity field against two available
mooring data records: KEO (the data are available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
and KESS mooring $\#$7 (the data are available online at \texttt{http://uskess.whoi.edu}).
For proper comparison, we use model data closest to these two moorings (west of
the region considered in the present study). Focusing on high-frequencies, we
calculate frequency spectral estimates every month. There are about 48 independent
spectral realizations.
